A meta-analysis of mental rotation ability in the first years of life

### 2022-04-18 code/analysis.Rmd
Reflect coding of gender mixed vs. split groups

```{r, compute_effects, echo=TRUE}
# Add columns to the table of included experiments
included %>%
  mutate(
    # Add difference between condition means
    mean_diff = case_when(
      !is.na(mean_diff) ~ mean_diff,
      TRUE ~ mean_novel - mean_familiar
    ),
    # Add d_z from paired t test of condition means (Rosenthal, 1991)
    d_z_t = t / sqrt(sample_size),
    # Add d_z from mean and standard deviation of the difference
    d_z_diff = mean_diff / sd_diff,
    # Add d_av from mean difference and standard devations (assumes r = 0)
    # (Cumming, 2012)
    sd_av = (sd_novel + sd_familiar) / 2,
    d_av = mean_diff / sd_av,
    # Add d from one-sample t test of novelty preference scores
    d_nov_pref = (nov_pref - 0.5) / sd_nov_pref,
    # Choose one type of outcome variable for each experiment
    yi = case_when(
      # 1. If d was reported directed
      !is.na(d) ~ d,
      # 2. If a paired sample t test was reported
      !is.na(d_z_t) ~ d_z_t,
      # 3. If the difference between means and its SD were reported
      !is.na(d_z_diff) ~ d_z_diff,
      # 4. If the individual condition means and their SDs were reported
      !is.na(d_av) ~ d_av,
      # 5. If a novelty preference score and its SD were reported
      !is.na(d_nov_pref) ~ d_nov_pref
    ),
    # Keep track which type of outcome measure was chosen for each article
    yi_type = case_when(
      !is.na(d) ~ "d",
      !is.na(d_z_t) ~ "d_z_t",
      !is.na(d_z_diff) ~ "d_z_diff",
      !is.na(d_av) ~ "d_av",
      !is.na(d_nov_pref) ~ "d_nov_pref",
      TRUE ~ "none"
    ) %>%
      factor(levels = c(
        "d", "d_z_t", "d_z_diff", "d_av", "d_nov_pref", "none"
      )),
    # Find studies with any value of d_z (from t test or mean / SD of the diff.)
    d_z = case_when(
      !is.na(d_z_t) ~ d_z_t,
      !is.na(d_z_diff) ~ d_z_diff
    ),
    # Compute SD of the difference based on d_z
    sd_z = mean_diff / d_z,
    # Compute correlation based on sd_z and condition SDs
    ri = (sd_z^2 - sd_novel^2 - sd_familiar^2) / (-2 * sd_novel * sd_familiar)
  ) %>%
  # Exclude experiments with redundant samples, missing stats, or age > 2 years
  filter(!redundant & !is.na(yi) & age_mean < 730) -> dat

# Overview of the different effect sizes
dat %>%
  select(
    article, group, yi_type, yi, d, d_z_t, d_z_diff, d_av, d_nov_pref, ri
  ) %>%
  print(n = Inf)

@@

```{r, compute_meta, echo=TRUE}
# Compute standard error of Cohen's d based on assumed correlation
# This will need a sensitivity analysis
r_assumed <- 0.5
dat %>%
  mutate(
    ni = sample_size,
    sei = sqrt(((2 * (1 - r_assumed)) / ni) + (yi^2 / (2 * ni))),
    age_months = age_mean / 30.417
  ) %>%
  select(article, group, age_months, female_percent, ni, yi, sei) %>%
  # Compute sampling variance from standard errors
  escalc(yi = yi, sei = sei, ni = ni, data = .) %>%
  tibble() -> dat_r

>>>>

```{r, compute_effects, echo=TRUE}
# Add columns to the table of included experiments
included %>%
  mutate(
    # Add difference between condition means
    mean_diff = case_when(
      !is.na(mean_diff) ~ mean_diff,
      TRUE ~ mean_novel - mean_familiar
    ),
    # Add d_z from paired t test of condition means (Rosenthal, 1991)
    d_z_t = t / sqrt(sample_size),
    # Add d_z from mean and standard deviation of the difference
    d_z_diff = mean_diff / sd_diff,
    # Add d_av from mean difference and standard devations (assumes r = 0)
    # (Cumming, 2012)
    sd_av = (sd_novel + sd_familiar) / 2,
    d_av = mean_diff / sd_av,
    # Add d from one-sample t test of novelty preference scores
    d_nov_pref = (nov_pref - 0.5) / sd_nov_pref,
    # Choose one type of outcome variable for each experiment
    yi = case_when(
      # 1. If d was reported directed
      !is.na(d) ~ d,
      # 2. If a paired sample t test was reported
      !is.na(d_z_t) ~ d_z_t,
      # 3. If the difference between means and its SD were reported
      !is.na(d_z_diff) ~ d_z_diff,
      # 4. If the individual condition means and their SDs were reported
      !is.na(d_av) ~ d_av,
      # 5. If a novelty preference score and its SD were reported
      !is.na(d_nov_pref) ~ d_nov_pref
    ),
    # Keep track which type of outcome measure was chosen for each article
    yi_type = case_when(
      !is.na(d) ~ "d",
      !is.na(d_z_t) ~ "d_z_t",
      !is.na(d_z_diff) ~ "d_z_diff",
      !is.na(d_av) ~ "d_av",
      !is.na(d_nov_pref) ~ "d_nov_pref",
      TRUE ~ "none"
    ) %>%
      factor(levels = c(
        "d", "d_z_t", "d_z_diff", "d_av", "d_nov_pref", "none"
      )),
    # Find studies with any value of d_z (from t test or mean / SD of the diff.)
    d_z = case_when(
      !is.na(d_z_t) ~ d_z_t,
      !is.na(d_z_diff) ~ d_z_diff
    ),
    # Compute SD of the difference based on d_z
    sd_z = mean_diff / d_z,
    # Compute correlation based on sd_z and condition SDs
    ri = (sd_z^2 - sd_novel^2 - sd_familiar^2) / (-2 * sd_novel * sd_familiar)
  ) -> dat

# Overview of the different effect sizes
dat %>%
  select(
    article,
    group,
    gender_split,
    yi_type,
    yi,
    d,
    d_z_t,
    d_z_diff,
    d_av,
    d_nov_pref,
    ri
  ) %>%
  print(n = Inf)

# Show number of experiments per gender grouping strategy
table(dat$gender_split)

# Extract mixed/split experiments only
dat_mixed <- filter(dat, gender_split %in% c("mixed", "mixed_only"))
dat_split <- filter(dat, gender_split %in% c("split", "split_only"))

# Extract both strategies combined (but excluding any redundant experiments)
dat_both <- filter(
  dat, gender_split %in% c("mixed", "mixed_only", "split_only")
)

@@

```{r, compute_meta, echo=TRUE}
# Compute standard error of Cohen's d based on assumed correlation
# This will need a sensitivity analysis
r_assumed <- 0.5
dat_both %>%
  mutate(
    ni = sample_size,
    sei = sqrt(((2 * (1 - r_assumed)) / ni) + (yi^2 / (2 * ni))),
    age_months = age_mean / 30.417
  ) %>%
  select(article, group, age_months, female_percent, ni, yi, sei) %>%
  # Compute sampling variance from standard errors
  escalc(yi = yi, sei = sei, ni = ni, data = .) %>%
  tibble() -> dat_r
